---
title: "Fragen an die flights data"
author: "Themenfeld SS 2021"
date: "04/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
data(flights)
flights = as.data.frame(flights)
```


#### how many flights started in January
solution 1 - via Boolean values:
```{r}
jan_flights = flights$month == 1
sum(jan_flights, na.rm = TRUE)
```
> $ only works with data frames and tibbles, not with "real" matrices

solution 2 - length()
```{r}
jan_flights = subset(flights, month == 1)
nrow(jan_flights) #number of rows
length(jan_flights$month)
```

solution 3 - my own subset
```{r}
jan_flights = flights[flights$month == 1, ]
```

solution 4 - all month
```{r}
table(flights$month)
```

#### number of delayed departures
```{r}
# Vectorized Operations
#DepDelay = flights[,"dep_time"] - flights[,"sched_dep_time"]
#Comparison yields Boolean values
delayedFlights = flights[,"dep_delay"] > 10
sum(delayedFlights, na.rm = TRUE)
mean(delayedFlights, na.rm = TRUE)
```

#### number of flights per month to ATL
```{r}
atl_flights = subset(flights, dest == "ATL")
table(atl_flights$month)

barplot(destATLmonth, main = "Flights to Atlanta per month", xlab = "month", ylab = "Num flights")
grid()
```

#### average delay on late arrivals
```{r}
mean_arr_delay <- mean(flights$arr_delay, na.rm= TRUE)

arr_del = flights[flights$arr_delay > 0]
mean(arr_del)

LateArrivals =subset(flights, arr_delay > 0) 
#mean_arr_delay <- mean(flights$arr_delay, na.rm= TRUE) #6.9
mean_arr_delay <- mean(LateArrivals$arr_delay, na.rm= TRUE) #40.34
```

#### distribution of departure delays for AA flights
```{r}
AAonly = flights$carrier == "AA"
AAonly = which(flights$carrier == "AA")
AAflights = flights[AAonly,]

AAdelays = flights[AAonly,"dep_delay"]
hAA = hist(AAdelays)
barplot(hAA$counts)
```
average delay
```{r}
Mean.EWR = mean(subset(LateArrivals,origin=="EWR")$arr_delay)
Mean.JFK = mean(subset(LateArrivals,origin=="JFK")$arr_delay)
Mean.LGA = mean(subset(LateArrivals,origin=="LGA")$arr_delay)
```
with aggregate()
```{r}
# 1. formula interface
aggregate(arr_delay ~ origin + month, data = LateArrivals, mean)
# 2. ...alternative
aggregate(LateArrivals$arr_delay,  list(LateArrivals$origin, LateArrivals$month), mean)
```

```{r}
library(dplyr)

by_tailnum <- group_by(flights, tailnum)
delay <- summarise(by_tailnum,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE))
delay <- filter(delay, count > 20, dist < 2000)

by_origin <- group_by(flights, origin)
delay <- summarise(by_origin,
                   count = n(),
                   delay = mean(arr_delay, na.rm = TRUE))

avg_arr_delay <- group_by(LateArrivals, origin)
arr_delayorig <- summarise(avg_arr_delay,
                           count = n(),
                           arr_delayorig = mean(arr_delay, na.rm = TRUE))
```

```{r}
DepDelay = flights[20,"dep_time"] - flights[20,"sched_dep_time"]
DepDelay
```


#### find the destination with the most flights
solution 1
```{r}
TableFluge = table(flights$dest)
tableflugesort = sort(TableFluge, decreasing = TRUE)
mostflights = tableflugesort[1]
mostflights
```
solution 2
```{r}
mostflights = table(flights$dest)
mostflights = sort(mostflights)
head(mostflights,1)
```