---
title: "Homework III, due May 27th"
author: "M Loecher"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(titanic)
data(titanic_train)
```

#### This homework is the absolute minimum extra work that you need to invest to keep up with the material in this course.

1. Review your understanding of **linear regression** using the book [An Introduction to Statistical Learning](http://faculty.marshall.usc.edu/gareth-james/ISL/)
  * Read chapters 3.1, 3.2.1, 3.3.1 
  * Reproduce Lab Exercises 3.6.4 (Interaction Terms)

```{r}
#
```

  * Reproduce Lab Exercises 3.6.6 (Qualitative Predictors)

```{r}
#
```

2. Load the `titanic` library and the `titanic_train` data. In the following we use `Survived` as the outcome (y):
   * fit a model containing the interactions between `Sex` and `Pclass`. Interpret the coefficients.
   * fit a model containing the interactions between `Sex` and `age`. Interpret the coefficients.
   * fit a model containing the interactions between `Sex`, `Pclass` and `age`. Interpret the coefficients.

```{r}
titanic_train$Pclass = factor(titanic_train$Pclass)
```

```{r}
fit1 = lm(Survived ~ Sex:Pclass, data = titanic_train)
summary(fit1)
```

```{r}
fit2.5 = lm(Survived ~ Sex+Age, data = titanic_train)
summary(fit2.5)
```

```{r}
fit2 = lm(Survived ~ Sex*Age, data = titanic_train)
summary(fit2)
```

```{r}
fit3 = lm(Survived ~ Sex:Pclass:Age, data = titanic_train)
summary(fit3)
```

## 6.7 The F test

### False Positives

create a model for testing

```{r}
# set.seed(321)

Nrows<-200; Ncols <- 20
    
ranData <- matrix(rnorm(Nrows*Ncols), ncol=Ncols)
colnames(ranData) <- c("Y", paste0("X", 1:(Ncols-1)))
ranData <- as.data.frame(ranData)
#build a model with ALL x variables
fitRan <- lm(Y ~ . , data = ranData)
sumRan <- summary(fitRan)
sumRan
```

rebuild the F-Statistic

```{r}
n <- 200
p <- 19
# z=2.3 What is the one-sided p value?
pVal <- 1-pnorm(2.3)

RSS <- sum(fitRan$res^2)
ESS <- sum((fitted(fitRan) - mean(ranData$Y))^2)
TSS <- RSS + ESS

Fstat = ((TSS-RSS)/p)/(RSS/(n-p-1))

# p value for F=1.114837
1-pf(Fstat, 19, 180)
```

***

# Chapter 8 Logistic Regression

```{r}
library(pracma)
x <- seq(-6, 6, length.out = 101)
y1 <- sigmoid(x)
plot(x, y1, type = "l", col = "darkblue", xlab = "", ylab = "", main = "Sigmoid Function(s)")
grid()

curve(1/(1+exp(-x)),-6,6,lwd=2,col="darkblue", main = "sigmoid function");grid()

curve(log(x/(1-x)),0,1,lwd=2,col="darkblue", main = "logit function");grid()
```

## 8.1 Interpretation of the coefficients

```{r}
titanic_train$Sex = factor(titanic_train$Sex)
titanic_train$Pclass = factor(titanic_train$Pclass)
naRows = is.na(titanic_train$Age)
NoMissingAge = titanic_train[!naRows,]
#logistic regression version:
model_logit = glm(Survived ~ Age + Sex + Pclass, data= NoMissingAge, family = binomial(link=logit))
summary(model_logit)$coef
```

lm() wird durch glm() ersetzt
ma
glm(..., family = binomial)