---
title: "Linear Regression"
subtitle: "How to create modles and test them"
author: "Themenfeld SS 2021"
date: "`r format(file.info(knitr::current_input())$atime, '%Y-%m-%d')`"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages = c(
  "dplyr",
#  "ggplot2",
  "wooldridge")

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

# 1   Loading and inspecting the data

```{r 1   Loading and inspecting the data}
data(hprice1)
```

## 1.1    A glimpse

```{r 1.1   A glimpse}
glimpse(hprice1)
```

## 1.2    Use the help function to inspect details of the dataset

```{r 1.2    Use the help function to inspect details of the dataset, eval=FALSE}
?hprice1
```

# 2   A first regression

```{r 2   A first regression}
my.lm1 <- lm(price ~ sqrft, data=hprice1)
summary(my.lm1)
```

## 2.1    Note:
We can define the formula separately

```{r 2.1    We can define the formula separately}
my.formula <- as.formula("price ~ sqrft")
my.lm2 <- lm(my.formula, data=hprice1)
```

### 2.1.1    Further hint

```{r 2.1.1    Further hint}
as.character(my.formula)
```

## 2.2    Using predict for predictions

```{r 2.2    Using predict for predictions}
my.price.hat <- predict(my.lm1)
head(my.price.hat)
```

# 3   Random data subsets
A simple way of obtaining a random sample is the following:

*First, reorder rows randomly  
*Second, take (for example) the first 80 percent of the rows  

```{r 3   Random data subsets}
my.random.index <- sample(nrow(hprice1), replace = FALSE)
head(my.random.index)
```

# 4    Tasks:

* Write a function my.RSS with two inputs: a string describing the formula and the dataset. The function carries out the regression and returns the RSS (residual sum of squares)  
* Write a function that has two input arguments: a data and a number for the split (e.g. 0.8 for 80 percent). The function returns randomly selected training and test datasets  

## 4.1   Task 1: Testing

```{r function definition my.RSS, echo=FALSE}
# two inputs: a string describing the formula and the dataset.
# carries out the regression and returns the RSS.
my.RSS <- function(my.str = "price ~ sqrft", my.data=hprice1){
  # convert formula string to formula
  tmp_formula <- as.formula(my.str)
  # carry out the regression
  tmp_lm <- lm(tmp_formula, data=my.data)
  tmp_rss <- sum((my.lm2$residuals)^2)
  return(tmp_rss)
}
```

```{r function call my.RSS}
my.RSS(my.str = "price ~ sqrft", my.data=hprice1)
```

## 4.2   Task 2: Testing 1

```{r function definition my.data.select1}
# two input arguments: a data and a number for the split.
# returns randomly selected training and test datasets
my.data.select1 <- function(my.data = hprice1, split=0.8){
  # absolute size of the sample
  sample_size <- round(split * nrow(my.data))
  # create a training dataset as a sample of the data
  training_data <- sample_n(my.data, sample_size)
  # indices of the samples used in training dataset
  sample_id <- as.numeric(rownames(training_data))
  # create a test dataset of the remainder of the data
  test_data <- my.data[-sample_id,]
  return(list(traindata = training_data,testdata = test_data))
}
```

```{r function definition my.data.select2}
# two input arguments: a data and a number for the split.
# returns randomly selected training and test datasets
my.data.select2 <- function(my.data = hprice1, split=0.8){
  # absolute size of the sample
  sample_size <- round(split * nrow(my.data))
  # create a training dataset as a sample of the data
  training_data <- my.data[sample(nrow(my.data),sample_size, replace = FALSE),]
  # indices of the samples used in training dataset
  sample_id <- as.numeric(rownames(training_data))
  # create a test dataset of the remainder of the data
  test_data <- my.data[-sample_id,]
  return(list(traindata = training_data,testdata = test_data))
}
```

```{r function call my.data.select1+2}
my.datasets1 <- my.data.select1(my.data = hprice1, split=0.8)
glimpse(my.datasets1)
my.datasets2 <- my.data.select2(my.data = hprice1, split=0.8)
glimpse(my.datasets2)
```

## 4.3   Task 2: Testing 2

```{r function call my.data.select1+2}
my.datasets1 <- my.data.select1(my.data = hprice1, split=0.8)
glimpse(my.datasets1)
my.datasets2 <- my.data.select2(my.data = hprice1, split=0.8)
glimpse(my.datasets2)
```


```{r predict.test}
my.lm2.1 <- lm(my.formula, data=my.datasets1$traindata)
summary(my.lm2.1)
my.lm2.2 <- lm(my.formula, data=my.datasets2$traindata)
summary(my.lm2.2)

predict.test1 <- predict(my.lm2.1, newdata = my.datasets1$testdata)
head(predict.test1)
predict.test2 <- predict(my.lm2.2, newdata = my.datasets2$testdata)
head(predict.test2)
```




```{r  function definition my.RSSTrainTest_, eval=FALSE, echo=FALSE}
my.RSSTrainTest = function(formula,my.data,split){
  # create Dataset
  fn.dataset = my.data.select(my.data = my.data, split = split)
  
  # get regression using traindata
  fn.lm = lm(formula, data = fn.dataset$traindata)
  
  # Use Regression to make predictions
  ## based on traindata
  predict.train = predict(fn.lm)
  ## based on testdata
  predict.test = predict(fn.lm,newdata = fn.dataset$traindata)
  

}
```

```{r  function definition my.RSSTrainTest}
my.RSSTrainTest <- function(my.str = "price ~ sqrft", my.data = hprice1, split = 0.8){
  tmp_datasets1 = my.data.select1(my.data = my.data, split = split)
  tmp_datasets2 = my.data.select2(my.data = my.data, split = split)
  
  tmp_rss_train1 <- my.RSS(my.str = my.str, my.data = my.datasets1$traindata)
  tmp_rss_test1 <- my.RSS(my.str = my.str, my.data = my.datasets1$testdata)
  tmp_rss_train2 <- my.RSS(my.str = my.str, my.data = my.datasets2$traindata)
  tmp_rss_test2 <- my.RSS(my.str = my.str, my.data = my.datasets2$testdata)
  
  print(paste("price1", my.datasets1$traindata$price[1]))
  print(paste("rss_train1 + rss_test1", tmp_rss_train1 + tmp_rss_test1, "rss_train2 + rss_test2", tmp_rss_train2 + tmp_rss_test2))
}
```


```{r function call my.RSSTrainTest}
my.RSSTrainTest("price ~ sqrft", my.data=hprice1, split=0.8)
```

```{r function call my.RSSTrainTest}
my.RSSTrainTest("price ~ sqrft", my.data=hprice1, split=0.8)
```